Some thoughts on Zermelo's iteration process in the context of the hyper2 package

### 2024-01-26 inst/podium.Rmd
consistentification
distinguish hankin2017 from hankin2017_rnw

---
title: "Does being on the podium affect Formula 1 drivers' performance in the next race?"
author: "R. K. S. Hankin"
output: bookdown::html_document2
bibliography: jss_hyper3.bib
---

@@

To cite the `hyper2` package in publications, please use @hankin2017a.
This short document defines the `podium()` function which tests the
hypothesis that a Formula 1 driver's being on the podium at one venue
affects his performance on the next.  Data from the 2007 season are
used.  It uses `hyper3` objects from the `hyper2` package.  A detailed
use-case for the top six drivers (Alonso, Hamilton, Heidfeld, Kubica,
Massa, and Raikkonen), and only the first five venues (Austria,
Malaysia, Bahrain, Spain, and Monaco) is given.

>>>>

---
title: "Does being on the podium affect Formula 1 drivers' performance in the next race?"
author: "R. K. S. Hankin"
output: bookdown::html_document2
bibliography: hyper2.bib  # copy vignettes/hyper2.bib to inst/
---

@@

To cite the `hyper2` package in publications, please use @hankin2017_rmd.
This short document defines the `podium()` function which tests the
hypothesis that a Formula 1 driver's being on the podium at one venue
affects his performance on the next.  Data from the 2007 season are
used.  It uses `hyper3` objects from the `hyper2` package.  A detailed
use-case for the top six drivers (Alonso, Hamilton, Heidfeld, Kubica,
Massa, and Raikkonen), and only the first five venues (Austria,
Malaysia, Bahrain, Spain, and Monaco) is given.

### 2024-01-26 inst/powerboat.Rmd
consistentification
distinguish hankin2017 from hankin2017_rnw

---
title: "powerboats with `hyper2`"
author: "R. K. S. Hankin"
output: bookdown::html_document2
bibliography: jss_hyper3.bib
---

@@

To cite the `hyper2` package in publications, please use @hankin2017a.
[powerboat](https://en.wikipedia.org/w/index.php?title=2018_F1_Powerboat_World_Championship&oldid=920386507)
racing is an important and prestigious sport.  Here I analyse a
dataset taken from the 2018 F1 Powerboat world championships.

>>>>

---
title: "powerboats with `hyper2`"
author: "R. K. S. Hankin"
output: bookdown::html_document2
bibliography: hyper2.bib  # copy vignettes/hyper2.bib to inst/
---

@@

To cite the `hyper2` package in publications, please use @hankin2017_rmd.
[powerboat](https://en.wikipedia.org/w/index.php?title=2018_F1_Powerboat_World_Championship&oldid=920386507)
racing is an important and prestigious sport.  Here I analyse a
dataset taken from the 2018 F1 Powerboat world championships.

### 2024-01-26 inst/pudding.Rmd
consistentification
distinguish hankin2017 from hankin2017_rnw

---
title: "Analysis of pudding following Davidson 1970"
author: "R. K. S. Hankin"
output: bookdown::html_document2
bibliography: jss_hyper3.bib
---

@@

To cite the `hyper2` package in publications, please use @hankin2017a.
This short document creates the `pudding` suite of R objects and
presents some preliminary analysis.  It is not particularly
well-structured or terse and is essentially a sequence of notes.  It
follows Davidson 1970 and the `PlackettLuce` package.

>>>>

---
title: "Analysis of pudding following Davidson 1970"
author: "R. K. S. Hankin"
output: bookdown::html_document2
bibliography: hyper2.bib  # copy vignettes/hyper2.bib to inst/
---

@@

To cite the `hyper2` package in publications, please use @hankin2017_rmd.
This short document creates the `pudding` suite of R objects and
presents some preliminary analysis.  It is not particularly
well-structured or terse and is essentially a sequence of notes.  It
follows Davidson 1970 and the `PlackettLuce` package.

### 2024-01-26 inst/race3.Rmd
distinguish hankin2017 from hankin2017_rnw

---
title: "Races with hyper3 objects"
author: "R. K. S. Hankin"
output: bookdown::html_document2
bibliography: hyper2.bib
---

@@

(takes about an hour to process without cache).  To cite the `hyper2`
package in publications, please use @hankin2017.  Consider a race in
which there are six runners 1-6 but we happen to know that three of
the runners (1,2,3) are clones of strength $p_a$, say, two of the
runners (4,5) have strength $p_b$, and the final runner (6) is of
strength $p_c$.  The runners race and the finishing order is:

>>>>

---
title: "Races with hyper3 objects"
author: "R. K. S. Hankin"
output: bookdown::html_document2
bibliography: hyper2.bib  # copy vignettes/hyper2.bib to inst/
---

@@

(takes about an hour to process without cache).  To cite the `hyper2`
package in publications, please use @hankin2017_rmd.  Consider a race in
which there are six runners 1-6 but we happen to know that three of
the runners (1,2,3) are clones of strength $p_a$, say, two of the
runners (4,5) have strength $p_b$, and the final runner (6) is of
strength $p_c$.  The runners race and the finishing order is:

### 2024-01-26 inst/retain.Rmd
consistentification
distinguish hankin2017 from hankin2017_rnw

---
title: "Keeping and discarding players: functions `keep()` and `keep_flawed()`"
author: "R. K. S. Hankin"
output: bookdown::html_document2
bibliography: jss_hyper3.bib
---

@@

To cite the `hyper2` package in publications, please use @hankin2017a.
Function `keep_flawed()` can be confusing.  Originally, the package
included a function `keep()` but this used flawed logic, hence the
name of its replacement.  Here is a tiny example that should clarify
things.  We will use a subset of `skating_table`:

>>>>

---
title: "Keeping and discarding players: functions `keep()` and `keep_flawed()`"
author: "R. K. S. Hankin"
output: bookdown::html_document2
bibliography: hyper2.bib  # copy vignettes/hyper2.bib to inst/
---

@@

To cite the `hyper2` package in publications, please use @hankin2017_rmd.
Function `keep_flawed()` can be confusing.  Originally, the package
included a function `keep()` but this used flawed logic, hence the
name of its replacement.  Here is a tiny example that should clarify
things.  We will use a subset of `skating_table`:

### 2024-01-26 inst/rock_paper_scissors_monster.Rmd
distinguish hankin2017 from hankin2017_rnw

---
title: "Quantifying the rock-paper-scissors effect"
author: "R. K. S. Hankin"
output: bookdown::html_document2
bibliography: hyper2.bib
---

@@

To cite the `hyper2` package in publications, please use @hankin2017;
takes about an hour to run without cache.  Here I use the `hyper2`
package to quantify the strength of a circular rock-paper-scissors
effect using the `hyper2` package, improving on previous results
[@west2008] which merely established its existence.  I use two
datasets, a three-player scoreline and a synthetic dataset referring
to wear on a seven-tooth gear wheel.

>>>>

---
title: "Quantifying the rock-paper-scissors effect"
author: "R. K. S. Hankin"
output: bookdown::html_document2
bibliography: hyper2.bib  # copy vignettes/hyper2.bib to inst/
---

@@

To cite the `hyper2` package in publications, please use @hankin2017_rmd;
takes about an hour to run without cache.  Here I use the `hyper2`
package to quantify the strength of a circular rock-paper-scissors
effect using the `hyper2` package, improving on previous results
[@west2008] which merely established its existence.  I use two
datasets, a three-player scoreline and a synthetic dataset referring
to wear on a seven-tooth gear wheel.

### 2024-01-26 inst/rowing.Rmd
consistentification
distinguish hankin2017 from hankin2017_rnw

---
title: "rowing: analysis of sculls in the 2016 Olympics"
author: "R. K. S. Hankin"
output: bookdown::html_document2
bibliography: jss_hyper3.bib
---

@@

To cite the `hyper2` package in publications, please use @hankin2017a.
This short document analyses the `rowing` dataset: sculls in the 2016
olympics.  It is not really standalone and is here as part of a
consistent dataset generation mechanism used in the `hyper2` package.
See the `hyper2` vignette for a more formal discussion.  In Olympic
sculling, up to six individual competitors race a small boat called a
scull over a course of length $2\,\mathrm{km}$; the object is to cross
the finishing line first.  Note that actual timings are irrelevant,
given the model, as the sufficient statistic is the order in which
competitors cross the finishing line.  The 2016 Summer Olympics is a
case in point: the gold and silver medallists finished less than $\sim
5$ milliseconds apart, corresponding to a lead of $\sim
2.5\,\mathrm{cm}$.  The package dataset is file `rowing.txt`, here are
the first three lines:

>>>>

---
title: "rowing: analysis of sculls in the 2016 Olympics"
author: "R. K. S. Hankin"
output: bookdown::html_document2
bibliography: hyper2.bib  # copy vignettes/hyper2.bib to inst/
---

@@

To cite the `hyper2` package in publications, please use @hankin2017_rmd.
This short document analyses the `rowing` dataset: sculls in the 2016
olympics.  It is not really standalone and is here as part of a
consistent dataset generation mechanism used in the `hyper2` package.
See the `hyper2` vignette for a more formal discussion.  In Olympic
sculling, up to six individual competitors race a small boat called a
scull over a course of length $2\,\mathrm{km}$; the object is to cross
the finishing line first.  Note that actual timings are irrelevant,
given the model, as the sufficient statistic is the order in which
competitors cross the finishing line.  The 2016 Summer Olympics is a
case in point: the gold and silver medallists finished less than $\sim
5$ milliseconds apart, corresponding to a lead of $\sim
2.5\,\mathrm{cm}$.  The package dataset is file `rowing.txt`, here are
the first three lines:

### 2024-01-26 inst/skating.Rmd
consistentification
distinguish hankin2017 from hankin2017_rnw

---
title: "Ladies' figure skating in the 2002 Winter Olympics"
author: "R. K. S. Hankin"
output: bookdown::html_document2
bibliography: jss_hyper3.bib
---

@@

To cite the `hyper2` package in publications, please use @hankin2017a.
This document analyses Bradley-Terry strengths of the competitors in
the Ladies' Figure Skating at the 2002 Winter Olympics.  In the
package, dataframe `skating_table` is copied from Lock and Lock (see
skating.Rd for more details).  It also creates file `data/skating.rda`
which contains R objects `skating`, `skating_table`, and
`skating_maxp`.

>>>>

---
title: "Ladies' figure skating in the 2002 Winter Olympics"
author: "R. K. S. Hankin"
output: bookdown::html_document2
bibliography: hyper2.bib  # copy vignettes/hyper2.bib to inst/
---

@@

To cite the `hyper2` package in publications, please use @hankin2017_rmd.
This document analyses Bradley-Terry strengths of the competitors in
the Ladies' Figure Skating at the 2002 Winter Olympics.  In the
package, dataframe `skating_table` is copied from Lock and Lock (see
skating.Rd for more details).  It also creates file `data/skating.rda`
which contains R objects `skating`, `skating_table`, and
`skating_maxp`.

### 2024-01-26 inst/skeleton.Rmd
consistentification
distinguish hankin2017 from hankin2017_rnw

---
title: "skeleton racing with the hyper2 package"
author: "R. K. S. Hankin"
output: bookdown::html_document2
bibliography: jss_hyper3.bib
---

@@

To cite the `hyper2` package in publications, please use @hankin2017a.
The [skeleton](https://en.wikipedia.org/wiki/Skeleton_(sport)) is a
winter sliding sport.  This short document discusses a dataset taken
from the [2018 Winter
Olympics](https://en.wikipedia.org/wiki/Skeleton_at_the_2018_Winter_Olympics_%E2%80%93_Men%27s)
and analyses the order statistic for the competitors.

>>>>

---
title: "skeleton racing with the hyper2 package"
author: "R. K. S. Hankin"
output: bookdown::html_document2
bibliography: hyper2.bib  # copy vignettes/hyper2.bib to inst/
---

@@

To cite the `hyper2` package in publications, please use @hankin2017_rmd.
The [skeleton](https://en.wikipedia.org/wiki/Skeleton_(sport)) is a
winter sliding sport.  This short document discusses a dataset taken
from the [2018 Winter
Olympics](https://en.wikipedia.org/wiki/Skeleton_at_the_2018_Winter_Olympics_%E2%80%93_Men%27s)
and analyses the order statistic for the competitors.

### 2024-01-26 inst/soling.Rmd
consistentification
distinguish hankin2017 from hankin2017_rnw

---
title: "soling"
author: "R. K. S. Hankin"
output: bookdown::html_document2
bibliography: jss_hyper3.bib
---

@@

```{r out.width='20%', out.extra='style="float:right; padding:10px"',echo=FALSE}
knitr::include_graphics(system.file("help/figures/hyper2.png", package = "hyper2"))
```

To cite the `hyper2` package in publications, please use @hankin2017a.
Soling is a type of boat raced in the 2000 Summer Olympics.  Here I
analyse race data taken from
[wikipedia](https://en.wikipedia.org/w/index.php?title=Sailing_at_the_2000_Summer_Olympics_\%E2\%80\%93_Soling&oldid=945362535)
using a Plackett-Luce extension of the Bradley-Terry model.

>>>>

---
title: "soling"
author: "R. K. S. Hankin"
output: bookdown::html_document2
bibliography: hyper2.bib  # copy vignettes/hyper2.bib to inst/
---

@@

```{r out.width='20%', out.extra='style="float:right; padding:10px"',echo=FALSE}
knitr::includme_graphics(system.file("help/figures/hyper2.png", package = "hyper2"))
```

To cite the `hyper2` package in publications, please use @hankin2017_rmd.
Soling is a type of boat raced in the 2000 Summer Olympics.  Here I
analyse race data taken from
[wikipedia](https://en.wikipedia.org/w/index.php?title=Sailing_at_the_2000_Summer_Olympics_\%E2\%80\%93_Soling&oldid=945362535)
using a Plackett-Luce extension of the Bradley-Terry model.

### 2024-01-26 inst/table_tennis_serve.Rmd
consistentification
distinguish hankin2017 from hankin2017_rnw

---
title: "The importance of serve in table tennis"
author: "R. K. S. Hankin"
output: bookdown::html_document2
bibliography: jss_hyper3.bib
---

@@

To cite the `hyper2` package in publications, please use @hankin2017a.
In this vignette, object `H` is created.  It is identical to object
`table_tennis` in the package; to see it type `data(table_tennis)` at
the command line, and to see the help page, type `help(table_tennis)`.
Suppose we have three players, "A","B", and "C" who play table tennis.
We seek to quantify the effect of the serve.  The scoreline is as
follows:

>>>>

---
title: "The importance of serve in table tennis"
author: "R. K. S. Hankin"
output: bookdown::html_document2
bibliography: hyper2.bib  # copy vignettes/hyper2.bib to inst/
---

@@

To cite the `hyper2` package in publications, please use @hankin2017_rmd.
In this vignette, object `H` is created.  It is identical to object
`table_tennis` in the package; to see it type `data(table_tennis)` at
the command line, and to see the help page, type `help(table_tennis)`.
Suppose we have three players, "A","B", and "C" who play table tennis.
We seek to quantify the effect of the serve.  The scoreline is as
follows:

### 2024-01-26 inst/tennis.Rmd
consistentification
distinguish hankin2017 from hankin2017_rnw

---
title: "Tennis doubles and ghosts"
author: "R. K. S. Hankin"
output: bookdown::html_document2
bibliography: jss_hyper3.bib
---

@@

To cite the `hyper2` package in publications, please use @hankin2017a.
This short document discusses a small dataset first presented by
@hankin2010.  There, results from tennis games among four players was
considered:

>>>>

---
title: "Tennis doubles and ghosts"
author: "R. K. S. Hankin"
output: bookdown::html_document2
bibliography: hyper2.bib  # copy vignettes/hyper2.bib to inst/
---

@@

To cite the `hyper2` package in publications, please use @hankin2017_rmd.
This short document discusses a small dataset first presented by
@hankin2010.  There, results from tennis games among four players was
considered:

### 2024-01-26 inst/ternaryplot_hyper2.Rmd
consistentification
distinguish hankin2017 from hankin2017_rnw

---
title: "Ternary plots for hyper2 samples"
author: "Robin Hankin"
output: html_document
bibliography: chess.bib
---

@@

To cite the `hyper2` package in publications please use @hankin2017.
Here I show how to use the excellent `Ternary` package [@smith2017] to
generate triangular plots for `hyper2` likelihood functions with size
3.  First load the `Ternary` and `hyper2` packages:

>>>>

---
title: "Ternary plots for hyper2 samples"
author: "Robin Hankin"
output: html_document
bibliography: hyper2.bib  # copy vignettes/hyper2.bib to inst/
---

@@

To cite the `hyper2` package in publications please use @hankin2017_rmd.
Here I show how to use the excellent `Ternary` package [@smith2017] to
generate triangular plots for `hyper2` likelihood functions with size
3.  First load the `Ternary` and `hyper2` packages:

### 2024-01-26 inst/volleyball.Rmd
consistentification
distinguish hankin2017 from hankin2017_rnw

---
title: "volleyball dataset"
author: "R. K. S. Hankin"
output: bookdown::html_document2
bibliography: jss_hyper3.bib
---

@@

To cite the `hyper2` package in publications, please use @hankin2017a.
This short document discusses a dataset first presented by @hankin2010,
although here only the first 52 observations are used.  A
volleyball _set_ is a Bernoulli trial between two disjoint subsets
of the players.  The two subsets are denoted (after the game) as the
_winners_ and the _losers_: these are denoted by `1` and `0`
respectively.

>>>>

---
title: "volleyball dataset"
author: "R. K. S. Hankin"
output: bookdown::html_document2
bibliography: hyper2.bib  # copy vignettes/hyper2.bib to inst/
---

@@

To cite the `hyper2` package in publications, please use @hankin2017_rmd.
This short document discusses a dataset first presented by @hankin2010,
although here only the first 52 observations are used.  A
volleyball _set_ is a Bernoulli trial between two disjoint subsets
of the players.  The two subsets are denoted (after the game) as the
_winners_ and the _losers_: these are denoted by `1` and `0`
respectively.

### 2024-01-26 inst/volvo.Rmd
consistentification
distinguish hankin2017 from hankin2017_rnw

---
title: "Race results from the twelfth edition of the
  round-the-world Volvo Ocean Race"
author: "Robin K. S. Hankin"
output: bookdown::html_document2
bibliography: jss_hyper3.bib
---

@@

To cite the `hyper2` package in publications, please use @hankin2017a.
This short document creates the `volvo` suite of R objects and
presents some preliminary analysis.  It refers to the 12th edition of
the Volvo Ocean Race, 2014-2015 in which seven racing teams competed
over nine legs.  Dataset taken from the wikipedia page (see references
below).

>>>>

---
title: "Race results from the twelfth edition of the
  round-the-world Volvo Ocean Race"
author: "Robin K. S. Hankin"
output: bookdown::html_document2
bibliography: hyper2.bib  # copy vignettes/hyper2.bib to inst/
---

@@

To cite the `hyper2` package in publications, please use @hankin2017_rmd.
This short document creates the `volvo` suite of R objects and
presents some preliminary analysis.  It refers to the 12th edition of
the Volvo Ocean Race, 2014-2015 in which seven racing teams competed
over nine legs.  Dataset taken from the wikipedia page (see references
below).

### 2024-01-26 inst/zermelo.Rmd
consistentification
distinguish hankin2017 from hankin2017_rnw

---
title: "Some thoughts on Zermelo's iteration process in the context of the hyper2 package"
author: "Robin K. S. Hankin"
output: bookdown::html_document2
bibliography: jss_hyper3.bib
---

@@

To cite the `hyper2` package in publications, please use @hankin2017a.
This short document sets out a use-case for functions `zermelo()` and
`pairwise()`.  First I create `n=10` competitors with Zipf strengths:

>>>>

---
title: "Some thoughts on Zermelo's iteration process in the context of the hyper2 package"
author: "Robin K. S. Hankin"
output: bookdown::html_document2
bibliography: hyper2.bib  # copy from vignettes/
---

@@

To cite the `hyper2` package in publications, please use @hankin2017_rmd.
This short document sets out a use-case for functions `zermelo()` and
`pairwise()`.  First I create `n=10` competitors with Zipf strengths:
